\mytitle

### 2015-08-05 MSc_diss.tex
New layout

\documentclass[11pt,a4paper]{article}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows
\renewcommand{\theequation}{R\arabic{equation}}
% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Blue},
    citecolor={Blue}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{epstopdf}
% Tables
\usepackage{multirow}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}

@@

\input{diss_titlepage}

@@

\begin{figure}[h]
\includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
\caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
\label{fig:O3observations}
\end{figure}

According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Table \ref{tab:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

\begin{table}[h] % O3 budget
\centering
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}
\label{tab:O3budget}
\begin{tabular}{ccl}
\hline
Burden, Tg & Production, Tg yr–1 & Reference \\
\hline
\multicolumn{3}{c}{Modelling studies} \\
$337\pm23$  & $4877\pm853$ & Young et al. (2013); ACCMIP \\
$323$	    & -	           & Archibald et al. (2011) \\
$330$	    & $4876$       & Kawase et al. (2011) \\
$312$       & $4289$       & Huijnen et al. (2010) \\
$334$       & $3826$       & Zeng et al. (2010) \\
$324$       & $4870$       & Wild and Palmer (2008) \\
$314$       & -	           & Zeng et al. (2008) \\
$319$	    & $4487$	   & Wu et al. (2007) \\
$372$	    & $5042$       & Horowitz (2006) \\
$349$	    & $4384$	   & Liao et al. (2006) \\
$344\pm39$	& $5110\pm606$ & Stevenson et al. (2006); ACCENT \\
$314\pm33$	& $4465\pm514$ & Wild (2007) (post-2000 studies) \\
\hline
\multicolumn{3}{c}{Observational studies} \\
$333$   	& -	           & Fortuin and Kelder (1998) \\
$327$	    & -	           & Logan (1999) \\
$325$	    & -	           & Ziemke et al. (2011); 60S–60N \\
$319–351$	& -	           & Osterman et al. (2008); 60S–60N \\
\hline
\end{tabular}
\end{table}

@@

In intermediate $NO_x$ conditions (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$), which subsequent photolysis leads to $O_3$ generation:
\begin{equation}\label{reac:HO2+NO=NO2+OH}
HO_2 + NO \rightarrow NO_2 + OH
\end{equation}
\begin{equation}\label{reac:CH3O2+NO=NO2+CH3O}
CH_3O_2 + NO \rightarrow NO_2 + CH_3O
\end{equation}
\begin{equation}\label{reac:NO2+hv=NO+O3P}
NO_2 + hv \rightarrow NO + O(^3P)
\end{equation}
\begin{equation}\label{reac:O3P+O2=O3}
!!!(R4)sameO(^3P) + O_2 \xrightarrow{M} O_3
\end{equation}
As shown in Figure ??? reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur. Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.

@@

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common. Its photolysis and further oxidation contributes to radical generation:
\begin{equation}\label{reac:HCHO+hv=H+HCO}
HCHO + hv \rightarrow H + HCO
\end{equation}
\begin{equation}\label{reac:HCO+O2=HO2+CO}
HCO + O_2 \rightarrow HO_2 + CO
\end{equation}
\begin{equation}\label{reac:H+O2=HO2}
!!!sameH + O_2 \xrightarrow{M} HO_2
\end{equation}
therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section. Their role is described in other sections as far as it is necessary.

@@

The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude. According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds. Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
\begin{equation}\label{eq:OPE}
OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
\end{equation}
where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate. $OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})). In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}. Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency. This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.

According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system. For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as: 
\begin{equation}\label{eq:epsilon}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
\end{equation}
where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
\begin{itemize}
\centering
\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
\end{itemize}
Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
\begin{itemize}
\centering
\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
\item[] $H + O_2 \xrightarrow{M} HO_2$
\item[] $horizontal line$
\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
\end{itemize}
or no peroxy radicals because of dry or wet deposition:
\begin{itemize}
\centering
\item[] $HCHO \xrightarrow{deposition}$
\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}\label{tab:ANmean} % mean AN mixing ratios
\begin{tabular}{cccc}
\hline
Environment     & Mean $\Sigma RONO_2$ & $\Sigma RONO_2$ breakdown & Reference \\
\hline
remote marine   & 10 pptv    & methyl                   & \citep{Reeves2007} \\
(North Atlantic)&            & ethyl                    & \\
airborne        &            & 1-propyl                 & \\   
                             && 2-propyl                & \\   
                             && 2-butyl                 & \\                                                                                               
                             && 2-methyl-3-butyl        & \\
                             && sum of 2- and 3-pentyl  & \\
\hline
rural           & 14 pptv    & methyl                   & \citep{Roberts1998} \\
(Nova Scotia, Canada)        && ethyl   & \\
ground                       && n-propyl & \\
                             && 2-propyl & \\
                             && 2-butyl  & \\ 
\hline
suburban                     & 17.48 pptv  & 2-propyl    & \citep{DeKock1994} \\
(Cape Retire, South Africa)  &(3.62-122.23)& 1-propyl & \\
ground                                    && 2-butyl  & \\
                                          && 1-butyl  & \\
                                          && 3-pentyl & \\
                                          && 2-pentyl & \\
                                          && 2-methyl-1-butyl & \\
                                          && i-pentyl & \\
\hline
suburban                     & 119.8 ppt  & $C_1$-$C_8$       & \citep{Flocke1998} \\
(Black Forest, Germany)      & (30-630 ppt) && \\ 
ground                                     &&& \\
\hline
polluted marine & approx 1 ppb & $NO_y$-($NO_x$+$HNO_3$+$PAN$) & \citep{Neuman2012} \\
(Gulf of Mexico)&(plume center) && \\
airborne &&& \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]\label{fig:Neu2008}
	\centering
	\begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT}.png}
        \caption{}
    \end{subfigure}%
    ~ % glues subplots on the same line somehow...
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
        \caption{}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
        \caption{}
    \end{subfigure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Chemical mechanism}\label{sec:method_chem}
The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: http://mcm.leeds.ac.uk/MCM. On a whole it includes 90 species and 165 reactions. The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates. These include:

>>>>

\documentclass[11pt,a4paper]{article}
% Font
\usepackage{helvet}
\renewcommand{\familydefault}{\sfdefault}
% Encoding and language
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
% Colors
\usepackage[usenames,dvipsnames]{color}
\usepackage{colortbl}
% Paper size and margins
%\usepackage{rotating}
\usepackage{pdflscape}
%\usepackage[a4paper,showframe]{geometry}
\usepackage{vmargin}
\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{1.5cm}
\usepackage{pdflscape}
\usepackage{afterpage}
% Indenting and line spacing
\linespread{1.5}
%\usepackage{setspace}
\usepackage{indentfirst}
%\setlength{\parindent}{1cm}
\setlength{\parskip}{0.5cm}
% Symbols
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
% Equations
\usepackage{eulervm} % upright font in eqns (Palatino)
\usepackage{cool}
\usepackage{mathtools} % for arrows

\newcounter{matriz}
\newenvironment{matriz}{\refstepcounter{matriz}\equation}{\tag{\thematriz}\endequation}
\renewcommand{\theequation}{R\arabic{equation}}

% References
\usepackage{natbib}
\bibliographystyle{apalike}
% Hyperlinks and labels
%\usepackage{showkeys}
\usepackage[linktocpage=true,plainpages=false,pdfpagelabels=false]{hyperref}
\definecolor{citecolor}{rgb}{0.2,0.7,0.9}
\definecolor{urlcolor}{rgb}{0,0,1}
\hypersetup{
    colorlinks, linkcolor={Blue},
    citecolor={Blue}, urlcolor={Black}
}
% Figures
\usepackage{subcaption}
\usepackage{alphalph}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usepackage{epstopdf}
\usepackage{wrapfig}
% Tables
\usepackage{multirow}
% Tikz styles
\tikzstyle{reg} = [rounded rectangle, text width=2cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{rad} = [rounded rectangle, draw=none, text width=2cm, fill=red!10, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]
\tikzstyle{long} = [rounded rectangle, draw=none, text width=3.5cm, fill=black!5, align=center, anchor=west,font=\sffamily\Large\bfseries, minimum height=1cm, inner sep=0]   
\tikzstyle{add} = [scale=1.5, draw=none,fill=none,align=center]
\tikzstyle{emp} = [draw=none,fill=none]
\tikzstyle{ar} = [->,draw=black!70,line width=2]
\tikzstyle{ln} = [-,draw=black!70,line width=2]
% Title etc
\newcommand{\mytitle}{Investigation of the relationship between tropospheric ozone production efficiency and carbon bond emissions}
\newcommand{\authorname}{Maria Zamyatina}
\newcommand{\authornumber}{Student No: 100106685}
\newcommand{\degname}{Master of Science}
\newcommand{\univname}{University of East Anglia}
\newcommand{\univaddr}{University Plain \\ Norwich NR47TJ UK}
\newcommand{\depname}{School of Environmental Sciences}
\newcommand{\mydate}{\the\year}

@@

\input{diss_titlepage}
%\onehalfspacing

@@

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{{./pics/Isaksen2009_O3}.png}
\caption{Observed surface ozone at different Northern Hemispheric surface stations \citep{Isaksen2009}.}
\label{fig:O3observations}
\end{figure}

According to the Intergovernmental Panel on Climate Change (IPCC) Fifth Assessment report (AR5) for present conditions (around year 2000) the global mean tropospheric ozone budget is approximately 331 Tg \citep{Myhre2013}. However, despite general agreement on how the drivers impact ozone concentrations at global and regional scales, its budget varies considerably between different modelling and observational studies \citep{Stevenson2006,Wild2007,Young2012}. Figure \ref{fig:O3budget} demonstrates that the range of the global tropospheric ozone budget estimates is approximately 60 Tg which is 18\% of the mean. In terms of ozone production the variation is even higher and accounts for 28\% of the corresponding value. In that regard, two questions arise: why does the tropospheric ozone budget differ between models and observations? And secondly, why does it vary to this extent, especially in terms of production? The only way to answer these questions is to perform a rigorous investigation of the factors that drive tropospheric ozone and attribute those differences to respective factors \citep{Young2012}.

\begin{figure}[t] % O3 budget
\caption{Summary of tropospheric ozone global budget model and observation estimates for present (about 2000) conditions. All uncertainties quoted as 1 standard deviation (68\% confidence interval). Adapted from \citep{Myhre2013}}
\centering
\label{tab:O3budget}
\includegraphics[width=0.7\linewidth]{o3_studies.eps}
\end{figure}

@@

In intermediate $NO_x$ conditions (rural areas of most industrialised countries), the peroxy radicals react with nitrogen monoxide ($NO$) to form nitrogen dioxide ($NO_2$), which subsequent photolysis leads to $O_3$ generation:
\begin{equation}\label{reac:HO2+NO=NO2+OH}
HO_2 + NO \rightarrow NO_2 + OH
\end{equation}
\begin{equation}\label{reac:CH3O2+NO=NO2+CH3O}
CH_3O_2 + NO \rightarrow NO_2 + CH_3O
\end{equation}
\begin{equation}\label{reac:NO2+hv=NO+O3P}
NO_2 + hv \rightarrow NO + O(^3P)
\end{equation}
\begin{equation}
O(^3P) + O_2 \xrightarrow{M} O_3\tag{\ref{reac:O3P+O2=O3}}
\end{equation}
As shown in Figure \ref{fig:O3scheme} reactions (\ref{reac:HO2+NO=NO2+OH}) and (\ref{reac:CH3O2+NO=NO2+CH3O}) form parts of the free-radical propagated $O_3$ forming cycles, which may occur a number of times before being halted by radical termination reactions (\ref{reac:HO2+HO2=H2O2+O2} and \ref{reac:HO2+CH3O2=CH3OOH+O2}), and the higher $NO_x$ level is the greater number of these cycles can occur. Therefore, in intermediate $NO_x$ conditions a net $O_3$ formation takes place and it is sensitive to $NO_x$.

\begin{figure}[h!]
\centering
\input{diss_NewO3scheme}
\caption{$O_3$ chemistry scheme}\label{fig:O3scheme}
\end{figure}

@@

The free-radical propagated oxidation of hydrocarbons leads to the generation of carbonyl compounds, among which formaldehyde ($HCHO$) is the most common. Its photolysis and further oxidation contributes to radical generation:
\begin{equation}\label{reac:HCHO+hv=H+HCO}
HCHO + hv \rightarrow H + HCO
\end{equation}
\begin{equation}\label{reac:HCO+O2=HO2+CO}
HCO + O_2 \rightarrow HO_2 + CO
\end{equation}
\begin{equation}
H + O_2 \xrightarrow{M} HO_2\tag{\ref{reac:H+O2=HO2}}
\end{equation}

therefore also having an impact on the rates of $O_3$ production \citep{Fowler2008} as do some other species, which are not mentioned in this section. Their role is described in other sections as far as it is necessary.

@@

The lifetimes of alkyl nitrates depend on the seasonality of the photolysis and $OH$ sink, which in turn vary with latitude and altitude. According to \citep{Clemitshaw1997} and \citep{Turberg1990} alkyl nitrates have atmospheric lifetimes of several days to several years, which makes them temporary reservoirs for the short lived $NO_x$ compounds. Being transported from polluted areas to remote background regions and eventually destroyed alkyl nitrates can 'release' $NO_2$ back to the atmosphere, thereby contributing to tropospheric ozone formation.
\subsection{Ozone, alkyl nitrates and carbon bonds}\label{sec:intro_CB}
In the previous sections we covered the chemistry that controls the production of $O_3$ in the troposphere. A traditional way to assess the efficiency of this production is to calculate the ratio:
\begin{matriz}\label{eq:OPE}
OPE = \dfrac{k_{HO_2+NO}[HO_2][NO]}{k_{OH+NO_2}[OH][NO_2]} = \dfrac{P_{O_3}}{L_{NO_x}}
\end{matriz}
where $OPE$ is the ozone production efficiency, $k_{HO_2+NO}$ - the rate coefficient of reaction (\ref{reac:HO2+NO=NO2+OH}), $k_{OH+NO_2}$ - the rate coefficient of reaction (\ref{reac:OH+NO2=HNO3}), $[X]$ - mixing ratios of corresponding species, $P_{O_3}$ – the $O_3$ production rate, $L_{NO_x}$ – the $NO_x$ loss rate. $OPE$ is determined by the amount of $HO_2$ reacting with $NO$ to form $NO_2$ and $OH$ (i.e., $k_{HO_2+NO}[HO_2][NO]$, (\ref{reac:HO2+NO=NO2+OH})) and the amount of $NO_x$ removed from the system by reaction of $OH$ with $NO_2$ to form nitric acid ($HNO_3$, reaction (\ref{reac:OH+NO2=HNO3})). In other words, $OPE$ shows the amount of ozone created ($P_{O_3}$) per molecule of $NO_x$ removed ($L_{NO_x}$) from the system \citep{Jacob1999}. Due to the reasons described in section \ref{sec:chem_mech} (comment at the bottom of the page!"No representation of the nitric acid ($HNO_3$) in the model"), we use another expression for calculations of the $O_3$ production efficiency. This expression has been recently proposed by Professor Matthew Evans (University of York) amid the mounting concerns about the need of thorough intercomparison of chemical mechanisms of global atmospheric models (including those used in the IPCC reports), why it is more universal and applicable to any kind of model.

According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system. For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as: 
\begin{matriz}\label{eq:epsilon}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
\end{matriz}
where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
\begin{itemize}
\centering
\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
\end{itemize}
Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
\begin{itemize}
\centering
\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
\item[] $H + O_2 \xrightarrow{M} HO_2$
\item[] $horizontal line$
\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
\end{itemize}
or no peroxy radicals because of dry or wet deposition:
\begin{itemize}
\centering
\item[] $HCHO \xrightarrow{deposition}$
\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\renewcommand{\arraystretch}{1}
\begin{table}[h!]
\caption{!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!}\label{tab:ANmean} % mean AN mixing ratios
\begin{tabular}{cccc}
\hline
Environment     & Mean $\Sigma RONO_2$ & $\Sigma RONO_2$ breakdown & Reference \\
\hline
remote marine   & 10 pptv    & methyl                   & \citep{Reeves2007} \\
(North Atlantic)&            & ethyl                    & \\
airborne        &            & 1-propyl                 & \\   
                             && 2-propyl                & \\   
                             && 2-butyl                 & \\                                                                                               
                             && 2-methyl-3-butyl        & \\
                             && sum of 2- and 3-pentyl  & \\
\hline
rural           & 14 pptv    & methyl                   & \citep{Roberts1998} \\
(Nova Scotia, Canada)        && ethyl   & \\
ground                       && n-propyl & \\
                             && 2-propyl & \\
                             && 2-butyl  & \\ 
\hline
suburban                     & 17.48 pptv  & 2-propyl    & \citep{DeKock1994} \\
(Cape Retire, South Africa)  &(3.62-122.23)& 1-propyl & \\
ground                                    && 2-butyl  & \\
                                          && 1-butyl  & \\
                                          && 3-pentyl & \\
                                          && 2-pentyl & \\
                                          && 2-methyl-1-butyl & \\
                                          && i-pentyl & \\
\hline
suburban                     & 119.8 ppt  & $C_1$-$C_8$       & \citep{Flocke1998} \\
(Black Forest, Germany)      & (30-630 ppt) && \\ 
ground                                     &&& \\
\hline
polluted marine & approx 1 ppb & $NO_y$-($NO_x$+$HNO_3$+$PAN$) & \citep{Neuman2012} \\
(Gulf of Mexico)&(plume center) && \\
airborne &&& \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[h!]
\centering
\includegraphics[width=0.5\linewidth]{{\picspath Neuman2012_Fig.4}.png}
\caption{Time series of reactive nitrogen species measured on 10 June 2010 in the MBL in plumes (a) 10 km and (b) 50 km downwind from DWH. $NO_x$, $NO_y$, and $HNO_3$ were measured once per s and $PAN$ was measured once per 2 s. $C_1$–$C_5$ alkyl nitrates were measured in whole air samples at locations indicated by the circles. The sum of alkyl nitrates (black lines) is inferred from the difference between $NO_y$ and ($NO_x$+$HNO_3$+$PAN$) \citep{Neuman2012}.}\label{fig:Neuman2012}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]\label{fig:Neu2008}
	\centering
	\begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.2_MeONO2_OCT_merge}.png}
        \caption{}
    \end{subfigure}%
    ~ % glues subplots on the same line somehow...
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3contol}.png}
        \caption{}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_NOxdiff}.png}
        \caption{}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{{\picspath Neu2008_Fig.3_OCT_O3diff}.png}
        \caption{}
    \end{subfigure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Chemical mechanism}\label{sec:method_chem}
The chemical mechanism of the model is based on recommendations from \citep{Atkinson2004} and information from the Master Chemical Mechanism, MCM v3.3 \citep{Jenkin1997,Saunders2003}, retrieved via website: \href{http://mcm.leeds.ac.uk/MCM}{http://mcm.leeds.ac.uk/MCM}. On a whole it includes 90 species and 165 reactions. The mechanism is developed to represent the chemistry that governs the production and loss of ozone, hydrocarbons and alkyl nitrates. These include:
